Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-22 tacl.tex
Added blurb for Japanese, plus other minor fixes

Work that uses collocation measures for multiword lexicon creation often assumes a small, fixed set of part-of-speech combinations, an assumption that breaks down entirely when considering the full range of formulaic sequences in a language. Relaxing this restriction, however, requires dealing with extreme overgeneration in terms of potential phrases; the main challenge is how to distinguish good FS in a huge collection of $n$-grams, many of which show basic statistical irregularities due to syntactic properties of the language or as the result of some relationship with a true FS. The former can be captured to some extent using a syntax-sensitive association measure like LPR, but addressing the latter requires a full, iterative model, where $n$-grams in essence complete with each other to form a compact, parsimonious formulaic vocabulary which best explain the statistics extracted from the corpus.

Our approach to FS identification involves optimization of the total explanatory power of a lattice where each node corresponds to an $n$-gram type. The explanatory power of the whole lattice is defined simply as a product of the \termdef{explainedness} of the individual nodes. Each node can be considered either ``on'' (is an FS) or ``off'' (is not an FS), and explainedness will be calculated differently depending on their status. Most fundamentally, $n$-gram nodes in the lattice are directionally connected to nodes consisting of $n+1$-grams which subsume them and $n-1$-grams which they subsume: that is, the (gapped) $n$-gram \ex{keep \gap under wraps} would, for instance be connected ``upwards'' to the node \ex{keep everything under wraps} and connected ``downwards'' to \ex{under wraps}. These directional relationships allow for two basic interactions between nodes in the lattice when a node is turned on: \termdef{covering}, which inhibits nodes below (subsumed by) a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model will tend not to choose \ex{under wraps} as an FS); and \termdef{clearing}, which inhibits nodes above a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model would avoid selecting \ex{keep everything under wraps} as an FS). A third, undirected mechanism is \termdef{overlapping}, where nodes inhibit each other due to overlaps in the corpus (\eg having both \ex{keep \gap under wraps} and \ex{be keep \gap under} as FS will be avoided). Given these relationships, we maximize the explanatory power of the entire lattice by iterating over the nodes and greedily optimizing the on/off choice relative to the local neighborhood of each node, until convergence.

@@

Once a set of relevant $n$-grams is identified (and counted), other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for some word $w_i$ in the context of a sequence of words of length $n$ which have a corresponding set of sequence of POS tags $t$, is given by:

\begin{displaymath}
LPR(w_i,w_{0,n}) = \max_{0 \leq j < k < n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}

@@

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (minLPR):

\begin{displaymath}
minLPR(w_{0,n}) = \min_{0 \leq i < n }{LPR(w_i,w_{0,n})}
\end{displaymath}

@@

We will influence these functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. Let's define $E_0$, then, as the minLPR based taken to the negative power of some as yet undefined function $d_0$:

\begin{displaymath}
E_0(t) = minLPR(t)^{-d_0(t)}
\end{displaymath}
And $E_1$ is the cost parameter $C$ taken to the negative power of an analogous function $d_1$:

\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not being influenced by other nodes. They are also restricted to always being positive.

@@

\termdef{Soft covering} is used in cases when a single $n$-gram does not entirely account for another, but a turned-on $n$-gram to some extent may explain some of the statistical irregularity of one lower in the lattice. For instance, in \figref{fig:example} \ex{keep \gap under} is not hard-covered by \ex{keep \gap under wraps} (since there are FS such as \ex{keep \gap under surveillance}, \ex{keep it under your hat}, \etc), but if \ex{keep \gap under wraps} is tagged as an FS, we nevertheless want to discount the portion of the \ex{keep \gap under} counts that corresponds to occurrences of \ex{keep \gap under wraps}. This is accomplished by increasing the turned-off explainedness of \ex{keep \gap under} (and thus making turning on less desirable) in the following manner: let $c(\dot)$ be the count function,and $s_1\dotts s_m$ be any turned-on nodes which are above $t$ in the lattice (covering nodes). Then, then the $d_0$ for a covered node $t$ is:

\begin{displaymath}
cover(t) = max(0,\frac{c(t) - \sum_{i=1}^{m}{c(s_i)}}{c(t)})
\end{displaymath}

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so explains nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which views shorter FS as helping to explain the presence of longer ones which contain them, beyond the FS-neutral syntactic explanation provided by minLPR. \termdef{Clearing} does this by changing the $E_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to explain covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are from the covering $n$-gram), in the reverse direction this logic fails. A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we increase $E_{0}$ of the higher node based on the ratio of the minLPR of two nodes, though only if the minLPR of the lower node is higher, which greatly restricts its effect. Generalized to consider the (rare) case of multiple clearing nodes, we define clear(t) as:

%in this paper, we consider two options: a hard clearing where any turned-on node will clear (set $C_{0}$ to 0 for) any node higher in the lattice which has a minLPR that is lower than it; and a soft

\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{minLPR(t)}{minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define $d_{0}(t)$ as:

\begin{displaymath}
d_0(t) = cover(t) \cdot clear(t)
\end{displaymath}

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. Though there are some exceptions, it is generally the case that independent formulaic sequences do not overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together, overlapping on the tokens \ex{keep \gap under}, we do not want both being selected as an FS, even in the case where both have high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than increasing the explainedness of having turned-off nodes, we decrease the explainedness of the turned-on nodes, which serves to express confusion at having overlapping FS. The idea behind covering and clearing is to encourage the model to turn on nodes which effectively explain the presence of nodes above and/or below it in the lattice, but for overlaps a penalty is a more appropriate mechanism. Let $oc(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1\dotts o_m$ refer to turned-on nodes which overlap with our target node $t$. We define $d_1$, the exponential term for the $E_1(t)$ as:

\begin{displaymath}
d_{1}(t) = \frac{c(t)}{c(t) - \sum_{i=1}^{m}{oc(o_i)}}
\end{displaymath}

@@

The other two languages we include in our evaluation are Croatian and Japanese. Relative to English, both languages are free-word order: we were interested in probing the challenges associated with using an $n$-gram approach to FS extraction in such languages. For Croatian, we used the X corpus, a collection of X. The POS regex for Croatian included XXX. .... For Japanese we used a portion of the X corpus of roughly the same size (by token) as the ICWSM. The POS regex for Japanese...

>>>>

Work that uses collocation measures for multiword lexicon creation often assumes a small, fixed set of part-of-speech combinations \cite{Evert:Krenn:2001,Pecina10}, an assumption that breaks down entirely when considering the full range of formulaic sequences in a language. Relaxing this restriction, however, requires dealing with extreme overgeneration in terms of potential phrases; the main challenge is how to distinguish good FS in a huge collection of $n$-grams, many of which show basic statistical irregularities due to syntactic properties of the language or as the result of some relationship with a true FS. The former can be captured to some extent using a syntax-sensitive association measure like LPR, but addressing the latter requires a full, iterative model, where $n$-grams in essence compete with each other to form a compact, parsimonious formulaic vocabulary which best explains the statistics extracted from the corpus.

Our approach to FS identification involves optimization of the total explanatory power of a lattice where each node corresponds to an $n$-gram type. The explanatory power of the whole lattice is defined simply as a product of the \termdef{explainedness} of the individual nodes. Each node can be considered either ``on'' (is an FS) or ``off'' (is not an FS), and explainedness will be calculated differently depending on this status. Most fundamentally, $n$-gram nodes in the lattice are directionally connected to nodes consisting of $(n+1)$-grams which subsume them and $(n-1)$-grams which they subsume: that is, the (gapped) $n$-gram \ex{keep \gap under wraps} would, for instance, be connected ``upwards'' to the node \ex{keep everything under wraps} and connected ``downwards'' to \ex{under wraps}. These directional relationships allow for two basic interactions between nodes in the lattice when a node is turned on: \termdef{covering}, which inhibits nodes below (subsumed by) a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model will tend not to choose \ex{under wraps} as an FS); and \termdef{clearing}, which inhibits nodes above a turned-on node (\eg, if \ex{keep \gap under wraps} is on, the model would avoid selecting \ex{keep everything under wraps} as an FS). A third, undirected mechanism is \termdef{overlapping}, where nodes inhibit each other due to overlaps in the corpus (\eg having both \ex{keep \gap under wraps} and \ex{be keep \gap under} as FS will be avoided). Given these relationships, we maximize the explanatory power of the entire lattice by iterating over the nodes and greedily optimizing the on/off choice relative to the local neighborhood of each node, until convergence.

@@

Once a set of relevant $n$-grams is identified (and counted), other statistics required to calculate \termdef{Lexical Predictability Ratios} (LPR) for each word in the $n$-gram are collected. LPR measures how predictable a word is in a lexical context as compared to how predictable it is given only syntactic context (over the same span of words). Formally, the LPR for some word $w_i$ in the context of a sequence of words of length $n$ which have a corresponding set of sequence of POS tags $t$, is given by:
\begin{displaymath}
LPR(w_i,w_{0,n}) = \max_{0 \leq j < k < n }{\frac{p(w_i|w_{j,k})}{p(w_i|t_{j,k})}}
\end{displaymath}

@@

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (minLPR):
\begin{displaymath}
minLPR(w_{0,n}) = \min_{0 \leq i < n }{LPR(w_i,w_{0,n})}
\end{displaymath}

@@

We will influence these functions by adding exponential terms, reflecting the fact we do not want to override the association measure, but rather modulate it. Let's define $E_0$, then, as the minLPR based taken to the negative power of some as yet undefined function $d_0$:
\begin{displaymath}
E_0(t) = minLPR(t)^{-d_0(t)}
\end{displaymath}
and $E_1$ is the cost parameter $C$ taken to the negative power of an analogous function $d_1$:
\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not being influenced by other nodes. They are also restricted to always being positive.

@@

\termdef{Soft covering} is used in cases when a single $n$-gram does not entirely account for another, but a turned-on $n$-gram to some extent may explain some of the statistical irregularity of one lower in the lattice. For instance, in \figref{fig:example} \ex{keep \gap under} is not hard-covered by \ex{keep \gap under wraps} (since there are FS such as \ex{keep \gap under surveillance}, \ex{keep it under your hat}, \etc), but if \ex{keep \gap under wraps} is tagged as an FS, we nevertheless want to discount the portion of the \ex{keep \gap under} counts that corresponds to occurrences of \ex{keep \gap under wraps}. This is accomplished by increasing the turned-off explainedness of \ex{keep \gap under} (and thus making turning on less desirable) in the following manner: let $c(\dot)$ be the count function,and $s_1\dotts s_m$ be any turned-on nodes which are above $t$ in the lattice (covering nodes). Then, then the $d_0$ for a covered node $t$ is:
\begin{displaymath}
cover(t) = \max(0,\frac{c(t) - \sum_{i=1}^{m}{c(s_i)}}{c(t)})
\end{displaymath}

@@

In general, covering prefers turning on longer, covering $n$-grams since doing so explains nodes lower in the lattice. Not surprisingly, it is generally desirable to have a mechanism working in opposition, \ie one which views shorter FS as helping to explain the presence of longer ones which contain them, beyond the FS-neutral syntactic explanation provided by minLPR. \termdef{Clearing} does this by changing the $E_0$ of nodes higher in the lattice when a lower node is turned-on. The basic mechanism is similar to covering, except that counts cannot be made use of in the same way; whereas it makes sense to explain covered nodes in proportion to the counts of their covering nodes (since the counts of the covered $n$-grams are from the covering $n$-gram), in the reverse direction this logic fails. A simple but effective solution which avoids introducing parameters makes use of the minLPR values of the relevant nodes. In the most common (two node) situation, we increase $E_{0}$ of the higher node based on the ratio of the minLPR of two nodes, though only if the minLPR of the lower node is higher, which greatly restricts its effect. Generalized to consider the (rare) case of multiple clearing nodes, we define clear(t) as:
%in this paper, we consider two options: a hard clearing where any turned-on node will clear (set $C_{0}$ to 0 for) any node higher in the lattice which has a minLPR that is lower than it; and a soft 
\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{minLPR(t)}{minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define $d_{0}(t)$ as:
\begin{displaymath}
d_0(t) = cover(t) \cdot clear(t)
\end{displaymath}

@@

The third mechanism of node interaction involves $n$-grams which overlap in the corpus. Though there are some exceptions, it is generally the case that independent formulaic sequences do not overlap. For example, given that \ex{be keep \gap under} and \ex{keep \gap under wraps} often appear together, overlapping on the tokens \ex{keep \gap under}, we do not want both being selected as an FS, even in the case where both have high minLPR. To address this problem, we use a mechanism somewhat different than above: rather than increasing the explainedness of having turned-off nodes, we decrease the explainedness of the turned-on nodes, which serves to express confusion at having overlapping FS. The idea behind covering and clearing is to encourage the model to turn on nodes which effectively explain the presence of nodes above and/or below it in the lattice, but for overlaps a penalty is a more appropriate mechanism. Let $oc(x,y)$ refer to number of times the $n$-grams corresponding to nodes $x$ and $y$ overlapped in the corpus, and $o_1\dotts o_m$ refer to turned-on nodes which overlap with our target node $t$. We define $d_1$, the exponential term for the $E_1(t)$ as:
\begin{displaymath}
d_{1}(t) = \frac{c(t)}{c(t) - \sum_{i=1}^{m}{oc(o_i)}}
\end{displaymath}

@@

The other two languages we include in our evaluation are Croatian and Japanese. Relative to English, both languages are free-word order: we were interested in probing the challenges associated with using an $n$-gram approach to FS extraction in such languages. For Croatian, we used the X corpus, a collection of X. The POS regex for Croatian included XXX. .... For Japanese, we used a subset of the 100M-page web corpus of \newcite{Shinzato+:2008} which was roughly the same size (in terms of token count) as the English corpus. We segmented and POS-tagged the corpus with MeCab \cite{Kudo:2008} using the UNIDIC morphological dictionary \cite{Den:2007}, and performed noun chunking based on a simple regex over the POS output.
% TJB: worth including the respective POS regexes in an appendix?
